Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-07-04 WRITEUP/ms.tex
some comment cleanup

\citestyle{aa}
\shorttitle{Dynamical Properties of
  % mm2 MC
              MCCs
  in Galaxies at the EoR}
\shortauthors{Leung et al.}

@@

\section{Introduction}
The growth of galaxies and their subsequent evolution are governed by the baryon cycle---galaxies accrete gas from the intergalactic medium (IGM) either directly from the cosmic web, or through mergers with other galaxies. This gas fuels \SF and feeds central
supermassive black holes. Subsequent feedback replenishes and enriches the circumgalactic medium by expelling some
part of this material.
Existing studies indicate that early galaxies are more gas-rich with molecular fractions higher than
those of present-day galaxies \citep[e.g.,][]{vandevoort11b, Decarli16a, Decarli17a}. 
Massive gas inflows from the IGM trigger gravitational instabilities that lead to the formation of molecular cloud complexes (MCCs)
that are typically more massive ($M_{\rm gas}\simeq 10^9$\,\Msun) and extended ($\simeq$ sub-kpc) than those observed in nearby galaxies (e.g., \citealt{Gabor13a, Hopkins14a, Inoue16a}).
Some theoretical works argue that the migration of such giant massive clumps are largely responsible for the buildup of the bulges of massive galaxies at redshift \z$\sim$\,0 \citep[e.g.,][]{Ceverino10a}.

%mm rearranged and combined this and the following paragraph to proceed from observation to conclusion
Early galaxies have higher star formation rates \citep[SFR; ][]{Behroozi13b, Sparre15a, Maiolino15a, Dunlop17a} and smaller sizes \citep[e.g.,][]{Bouwens11a, Ono13a} compared to the local population \citep[see also a review by][]{Stark16a}.
%
As a consequence, we expect them to be significantly more ionized, and have intense and hard interstellar radiation fields (ISRF). Since their metallicity and dust content are also expected to be lower in these early evolutionary stages, shielding of UV photons---responsible for the photoheating of the gas---  is strongly reduced. Such differences in turn affect the regulation of the thermal and chemical state of the multi-phase interstellar medium (ISM). Studying ISM properties of early galaxies is essential for understanding how \SF proceeds under more extreme conditions.
%
Even in the local Universe, where detailed \obs can be made, variations in molecular cloud
%       \MM{do Hughes et al. study cloud complexes, as you do, or can you derive info about complexes from their data?}
%     \DL{This is just a statement to say there are variations among local GMCs. Together with the different galaxy properties found for early galaxies, it is interesting to study properties of molecular structures at earlier epochs. }
properties have been observed between different galaxy populations (see e.g., \citealt{Hughes10a, Hughes13b}).  Given that \highz galaxies statistically represent the early evolutionary stages of present-day galaxies, it is thus reasonable to pose the question: 
{\it what are the physical properties of
molecular cloud structures in early galaxies, and how do they differ from those found in local \galpop?}
%

FIR fine-structure lines (e.g., \cii, \nii, and \oiii), and CO and [\ci]~lines are key diagnostics for constraining the ISM conditions of galaxies. They also provide highly complementary information on different ISM phases (ionized, atomic, molecular; e.g., \citealt{Scoville74a, Rubin85a, Malhotra01a}).
%
Global measurements of these diagnostics in \highz galaxies have provided preliminary information on their global properties (e.g., gas masses, gas temperature, and radiation field intensity). However, spatially resolving their ISM is necessary to fully understand many aspects of galaxy evolution and the physics behind their intense \SF (SFR\ssim100$-$3000\,\Msun\,yr\pmOne; 
see e.g., the review by \citealt{CW13}).
%
To date, spatially resolved ISM properties of \highz galaxies have only been mapped in a handful of 
galaxies at high redshift using tracers such as dust continuum, CO, and \cii lines (e.g., \citealt{Swinbank11a, Hodge15a, Ferkinhoff15a, 
Hodge16a, Leung19a}). These studies find that galaxies close to the
peak of cosmic \SF ($z$\ssim2) are more molecular gas-rich, turbulent,
and clumpy than nearby galaxies,
% mm2
    although luminous compact blue galaxies may represent local
    analogs of these gas-rich, clumpy systems \citep{Garland15}.

Earlier epochs still represent an essentially uncharted territory for ISM investigations. At present, it remains unclear how \SF proceeds in the (sub-)$L^*$ galaxy population at \z$\gtrsim$\,6  which is responsible for producing the bulk of the UV photons that reionized the Universe.
%
While ALMA has enabled the detection of \cii158\,$\micron$ and CO line
emission in normal (SFR$<$\,100\,\Msun\,yr\pmOne) galaxies at \z$>$\,6
over the past few years \citep[e.g.,][]{Carniani18b, Odorico18a}, the
first spatially resolved observations are just starting to become
available \citep[e.g., ][]{Jones17a}.
% mm2
    On the theoretical side, a schematic investigation by
    \citet{Behrendt16} revealed that kiloparsec-scale clumps likely consist
    of many smaller clumps formed by gravitational instability, a hierarchical
    structure that persists in an idealized disk model by \citet{Behrendt19}.

%mm To answer these questions, isolate key issues, and  accompany ongoing/future experimental efforts,
To understand the physical properties of MCCs in early galaxies,
we have undertaken a detailed numerical study whose aim is to characterize the dynamical properties of the star-forming %mm molecular cloud complexes
MCCs
in prototypical (i.e., $L^*$) galaxies in the Epoch of Reionization (EoR).

@@

\section{Physical Background}\label{sec:Back}
%mm2 [revised paragraph]
We begin by introducing some empirical relations commonly mentioned in
observational studies of molecular clouds in the literature, as well
as describing the gravitational instability of galactic disks, which
might be driving MCC formation. These concepts will be used in the subsequent analysis of our simulations.

\subsection{Larson's Relations and Virial Parameter}\label{sec:PVE}
\citet{Larson81a} discussed a number of relations among Galactic
molecular cloud
properties, namely the linewidth-size, density-size, and mass-size relations. Larson relations are routinely used for comparing properties of molecular structures in different galactic environments. They also represent a useful framework to analyze our results as they
have been argued to arise from the interplay between gravity and
% mm2 turbulence encoded in the
    velocity dispersion given by
virial theorem
(note, however, the alternative interpretation involving gravitational collapse by
e.g., \citealt{Ballesteros-Paredes11a}).

The virial theorem for a distribution of unmagnetized gas can be written as \citep{McKee92a}
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the
Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of
the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and
$\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm gas}$, radius $R$, and root-mean-square velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions.
%mm Neglecting magnetic field and
Defining $P_{\rm ext}$ as the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the right-hand side represent the kinetic energy, external pressure, and gravitational potential energy terms.

@@

Based on Equation~\ref{eqn:SVE}, a one-to-one mapping between
$\sigma_{\rm gas}^2/R$ and $\Sigma$ is therefore expected for a
virialized cloud, since $\sigma_{\rm
  gas}^2/R\propto\Sigma$.
% mm2
    \citep{Heyer09a} pointed out that the original size-linewidth
    relation implies constant surface density in this interpretation.
    The limited dynamic range in column density of CO observations
    appears to account for this \citep[see also][]{Ballesteros-Paredes11a}.
Deviations from this relation are
% mm2 usually
        often
attributed to a significant contribution from external pressure as per
Equation~\ref{eqn:v0} (see e.g., \citealt{Heyer09a, Hughes10a,
  Hughes13b, Meidt13a}). Similar conclusions are also reached from the
analysis of clouds forming in a Milky-Way like galaxy simulated with a $\simeq 4\,{\rm pc}$ resolution \citep[i.e.][]{grisdale:2018}.

@@

%\subsection{Gravitational Instabilities}\label{sec:Q}
\subsection{Toomre stability analysis}\label{sec:Q}

@@

From the dispersion relation, a parameter $Q$ can be derived such that $Q < 1$ when instability occurs, that reproduces this inequality to order unity. For a collisionless fluid---such as an ensemble of stars--- this parameter is \citep{Toomre64a}
\begin{equation}
Q_{\star} \equiv\frac{\sigma_{\star}\kappa}{3.36 G \Sigma_{\star}}\,.
\end{equation}
The equivalent parameter for a collisional gas was derived by \citet{Goldreich65a}
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma_{\rm gas}\kappa}{\pi G \Sigma_{\rm gas}}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), instability occurs on scales $k$ such that $Q < Q_{\rm crit}\simeq1$ (or equivalently $\omega^2 < 0$ in \Eq{3Ddisp}). A frequently used observable proxy for $Q$ is the ratio of disk circular velocity to root-mean-square velocity dispersion $v_{\rm circ}/\sigma_{\rm disk}$ \citep[e.g.,][]{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}.
% https://ned.ipac.caltech.edu/level5/March15/Glazebrook/Glazebrook5.html
%

In our stability analysis, we account for the combined effect of gas and stars \citep[derived exactly by][]{Rafikov01a}, and for the non-negligible disk thickness. This is done by adopting an approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modifies the $Q$ parameter for gas and stars by accounting for the vertical velocity dispersion
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the finite disk thickness reduces the gravity in the vertical direction, thereby making it easier for a system to maintain stability, and thus lowering the critical Toomre $Q_{\rm crit}$ from $\simeq$\,1 to 0.67 \citep{Goldreich65a}.
%
On the other hand, including the contribution of the stellar component promotes gravitational instability, and thus increases
$Q_{\rm crit}$, more-so if the stars have low velocity dispersion.
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

% ------------------
\section{Numerical Simulations}\label{sec:sim}
The simulations used in this work are described by \citet{Pallottini17a, Pallottini17b} and are briefly summarized here.
%
\ncode{Serra}\footnote{Greenhouse in Italian.} is a suite of cosmological zoom-in simulations performed using Eulerian hydrodynamics and adaptive mesh refinement (AMR) techniques to achieve high spatial resolution in regions of interest (e.g., regions of high density).
%
In particular, it uses a modified version of \ncode{ramses} \citep{Teyssier02a} as the AMR backend. The simulation used here covers a comoving box of 20\,Mpc $h$\pmOne in size. The simulation zooms in on a target halo of mass $M_{\rm DM}\simeq10^{11} \Msun$. The Lagrangian region of the halo (2.1\,Mpc $h$\pmOne) has a dark matter mass resolution of $\simeq 6\times 10^4 \Msun$, equivalent at initial density to a baryonic mass resolution of $1.2 \times 10^4 \Msun$. This region is spatially refined with a quasi-Lagrangian criterion based on a mass threshold, so that a cell is refined if its total (dark+baryonic) mass exceeds the mass resolution by a factor of eight. The finest refined cell allowed in the zoom in region has size $l_{\rm cell}\simeq$\,30\,pc (at $z = 6$), i.e. sizes are comparable to the sizes of local giant molecular clouds \citep[e.g.,][]{Sanders85a, Federrath13a, Goodman14a},
The models include a non-equilibrium chemical network following e$^{-}$, H$^+$, H$^-$, He, He$^+$, He$^{++}$, H$_2$, and H$_2^+$ \citep{Grassi14a,Bovino16a}. Of particular importance here is that abundances are calculated using an on-the-fly non-equilibrium formation of molecular hydrogen scheme described by \citet{Pallottini17a}.

Star formation is modeled using an H$_2$-based prescription of the
Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar
tracks from \ncode{starburst99} and include stellar feedback from
supernovae (SNe) as well as winds from OB and asymptotic giant branch
stars. Coupling to the gas is implemented via a sub-grid model for
blastwaves, which accounts for radiative energy losses inside the
cell. The remaining energy is injected into the ISM in
%mm2 kinetic and thermal energy forms. Via feedback processes the gas
%can acquire energy in
both thermal and non-thermal form \citep{Agertz13a}; the latter is
a term that mimics unresolved turbulence. The non-thermal energy is
not affected by radiative cooling but rather is dissipated on the eddy
turn-over time scale \citep{Maclow99a}. See \citet{Pallottini17b} for
further details.
%mm2 [moved down to unify discussion of non-thermal energy]
Radiation pressure on dust and gas is also included. Photoionization from local sources is neglected in this simulation \citep[see][for its effect]{Pallottini19a, Decataldo19a}.

@@

One of the main advantages of studying galaxies in simulations is that we can examine how their dynamical properties evolve with time, which
is interesting in order to understand the physical processes that determine the morphology and dynamical properties of galaxies, which in turn affect their \SF.
%mm2 [the first three lines of the following sentence were commented
%out, but not the last line.  Since it makes sense, and the following sentence starts with
%Meanwhile, I have recovered the entire sentence.] 
This is advantageous especially at early cosmic epochs, when the
densest structures are beginning to form; gas is constantly being
accreted onto the central galaxy from the cosmic web and 
%mm2  \section{•} atellite
             satellite
galaxies, thereby leading to bursts of \SF. Meanwhile, tidal forces resulting from interactions with these surrounding galaxies can disrupt the main disk and arms, likely leading to different dynamical states for the molecular structures compared to more evolved galaxies found at a later cosmic time (e.g., some molecular structures may disperse while others may agglomerate into more massive ones). % mainly expecting differences in alpha_vir, sigma, M_cl.
%mm [delete para break]
The \SF history\footnote{Note that here the SFR is calculated based on the stellar mass formed in the last 4 Myr within 3.5 kpc from the galaxy center of mass. The SFR plotted in Figure 2 of \citet{Pallottini17b} is a factor of two higher since there the SFR accounts for the contribution from massive satellite galaxies within the virial radius ($\approx$15\,kpc).} of \flower is shown in \Fig{SFH}. The SFR of \flower varies between $\sim$30--80 \Msun\,yr\pmOne as it evolves from an actively accreting phase to a starburst phase after a merger, and then back to a relatively quiescent phase, over the simulated $\approx 700$\,Myr.

% mm2 [move this sentence up to give sections in order]
In \Sec{dist}, we show the importance of rotation support from large-scale motions in the MCC dynamics 
by comparing their properties in the disturbed phase of \flower and in the disk-like ordered phase.
%
Given the stochastic nature of \flower in its star formation history, we mostly focus on few of its most extreme evolutionary stages in this work (see \Sec{singless}).
%
These phases correspond to (a) an intensely accreting phase and (b) a starburst phase (\Fig{SFH}). We are interested in determining whether MCC properties are sensitive to these different dynamical conditions. For completeness, we also show relationships of MCC properties examined for other evolutionary stages traced in the simulation (see \Sec{ncut}).

@@

In \Fig{h2density}, we show the distribution of H$_2$ number density $n_{\rm H2}$ for \flower during its accreting phase, including the contribution from the gas within 3.5 kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of densities where clumps are found based on morphological analysis\footnote{For $n_{\rm H2}\gtrsim1$\,\cc, the fourth Minkowski functional of the H$_{2}$ density field is significantly larger than zero. This implies that the field is made of isolated components. See Fig. 6 in \citet{Pallottini17b}.} \citep{Pallottini17b}.
%
In each evolutionary stage, we identify MCCs by applying density cuts to the {\em H$_2$ density} distribution, i.e., MCCs
are selected at $n_{\rm cut} \leq n_{\rm H2}$. We select 10 equally-spaced cuts in H$_2$ density in log scale: ${(n_{\rm cut}/{1 {\rm cm}^{-3})}}\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 18.96, 31.62]$.
% We also tested variations in $n_{\rm cut}$ range and found no qualitative differences (see \Sec{ncut}).
Note that with these choices we are including MCCs that are not fully molecular ($n = 0.5 n_{\rm H2}$).

>>>>

\citestyle{aa}
\shorttitle{Dynamical Properties of MCCs in Galaxies at the EoR}
\shortauthors{Leung et al.}

@@

\section{Introduction}
The growth of galaxies and their subsequent evolution are governed by the baryon cycle---galaxies accrete gas from the intergalactic medium (IGM) either directly from the cosmic web, or through mergers with other galaxies. This gas fuels \SF and feeds central supermassive black holes. Subsequent feedback replenishes and enriches the circumgalactic medium by expelling some
part of this material.
%
Existing studies indicate that early galaxies are more gas-rich with molecular fractions higher than those of present-day galaxies \citep[e.g.,][]{vandevoort11b, Decarli16a, Decarli17a}. Massive gas inflows from the IGM trigger gravitational instabilities that lead to the formation of molecular cloud complexes (MCCs) that are typically more massive ($M_{\rm gas}\simeq 10^9$\,\Msun) and extended ($\simeq$ sub-kpc) than those observed in nearby galaxies (e.g., \citealt{Gabor13a, Hopkins14a, Inoue16a}). Some theoretical works argue that the migration of such giant massive clumps are largely responsible for the buildup of the bulges of massive galaxies at redshift \z$\sim$\,0 \citep[e.g.,][]{Ceverino10a}.

Early galaxies have higher star formation rates \citep[SFR; ][]{Behroozi13b, Sparre15a, Maiolino15a, Dunlop17a} and smaller sizes \citep[e.g.,][]{Bouwens11a, Ono13a} compared to the local population \citep[see also a review by][]{Stark16a}.
%
As a consequence, we expect them to be significantly more ionized, and have intense and hard interstellar radiation fields (ISRF). Since their metallicity and dust content are also expected to be lower in these early evolutionary stages, shielding of UV photons---responsible for the photoheating of the gas---  is strongly reduced. Such differences in turn affect the regulation of the thermal and chemical state of the multi-phase interstellar medium (ISM). Studying ISM properties of early galaxies is essential for understanding how \SF proceeds under more extreme conditions.
%
Even in the local Universe, where detailed \obs can be made, variations in molecular cloud properties have been observed between different galaxy populations (see e.g., \citealt{Hughes10a, Hughes13b}).  Given that \highz galaxies statistically represent the early evolutionary stages of present-day galaxies, it is thus reasonable to pose the question: {\it what are the physical properties of molecular cloud structures in early galaxies, and how do they differ from those found in local \galpop?}
%

FIR fine-structure lines (e.g., \cii, \nii, and \oiii), and CO and [\ci]~lines are key diagnostics for constraining the ISM conditions of galaxies. They also provide highly complementary information on different ISM phases (ionized, atomic, molecular; e.g., \citealt{Scoville74a, Rubin85a, Malhotra01a}).
%
Global measurements of these diagnostics in \highz galaxies have provided preliminary information on their global properties (e.g., gas masses, gas temperature, and radiation field intensity). However, spatially resolving their ISM is necessary to fully understand many aspects of galaxy evolution and the physics behind their intense \SF (SFR\ssim100$-$3000\,\Msun\,yr\pmOne; see e.g., the review by \citealt{CW13}).
%
To date, spatially resolved ISM properties of \highz galaxies have only been mapped in a handful of galaxies at high redshift using tracers such as dust continuum, CO, and \cii lines (e.g., \citealt{Swinbank11a, Hodge15a, Ferkinhoff15a, Hodge16a, Leung19a}). These studies find that galaxies close to the peak of cosmic \SF ($z$\ssim2) are more molecular gas-rich, turbulent, and clumpy than nearby galaxies, although luminous compact blue galaxies may represent local analogs of these gas-rich, clumpy systems \citep{Garland15}.

Earlier epochs still represent an essentially uncharted territory for ISM investigations. At present, it remains unclear how \SF proceeds in the (sub-)$L^*$ galaxy population at \z$\gtrsim$\,6  which is responsible for producing the bulk of the UV photons that reionized the Universe.
%
While ALMA has enabled the detection of \cii158\,$\micron$ and CO line emission in normal (SFR$<$\,100\,\Msun\,yr\pmOne) galaxies at \z$>$\,6 over the past few years \citep[e.g.,][]{Carniani18b, Odorico18a}, the first spatially resolved observations are just starting to become available \citep[e.g., ][]{Jones17a}.
%
On the theoretical side, a schematic investigation by \citet{Behrendt16} revealed that kiloparsec-scale clumps likely consist of many smaller clumps formed by gravitational instability, a hierarchical structure that persists in an idealized disk model by \citet{Behrendt19}.

To understand the physical properties of MCCs in early galaxies, we have undertaken a detailed numerical study whose aim is to characterize the dynamical properties of the star-forming MCCs in prototypical (i.e., $L^*$) galaxies in the Epoch of Reionization (EoR).

@@

\section{Physical Background}\label{sec:Back}
We begin by introducing some empirical relations commonly mentioned in observational studies of molecular clouds in the literature, as well as describing the gravitational instability of galactic disks, which might be driving MCC formation. These concepts will be used in the subsequent analysis of our simulations.

\subsection{Larson's Relations and Virial Parameter}\label{sec:PVE}

\citet{Larson81a} discussed a number of relations among Galactic molecular cloud properties, namely the linewidth-size, density-size, and mass-size relations. Larson relations are routinely used for comparing properties of molecular structures in different galactic environments. They also represent a useful framework to analyze our results as they have been argued to arise from the interplay between gravity and velocity dispersion given by virial theorem (note, however, the alternative interpretation involving gravitational collapse by e.g., \citealt{Ballesteros-Paredes11a}).

The virial theorem for a distribution of unmagnetized gas can be written as \citep{McKee92a}
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the
Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of
the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and
$\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm gas}$, radius $R$, and root-mean-square velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions. Defining $P_{\rm ext}$ as the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the right-hand side represent the kinetic energy, external pressure, and gravitational potential energy terms.

@@

Based on Equation~\ref{eqn:SVE}, a one-to-one mapping between $\sigma_{\rm gas}^2/R$ and $\Sigma$ is therefore expected for a virialized cloud, since $\sigma_{\rm gas}^2/R\propto\Sigma$.
%
\citep{Heyer09a} pointed out that the original size-linewidth relation implies constant surface density in this interpretation. The limited dynamic range in column density of CO observations appears to account for this \citep[see also][]{Ballesteros-Paredes11a}.
%
Deviations from this relation are often attributed to a significant contribution from external pressure as per Equation~\ref{eqn:v0} (see e.g., \citealt{Heyer09a, Hughes10a, Hughes13b, Meidt13a}). Similar conclusions are also reached from the
analysis of clouds forming in a Milky-Way like galaxy simulated with a $\simeq 4\,{\rm pc}$ resolution \citep[i.e.][]{grisdale:2018}.

@@

\subsection{Toomre stability analysis}\label{sec:Q}

@@

From the dispersion relation, a parameter $Q$ can be derived such that $Q < 1$ when instability occurs, that reproduces this inequality to order unity. For a collisionless fluid---such as an ensemble of stars--- this parameter is \citep{Toomre64a}
\begin{equation}
Q_{\star} \equiv\frac{\sigma_{\star}\kappa}{3.36 G \Sigma_{\star}}\,.
\end{equation}
The equivalent parameter for a collisional gas was derived by \citet{Goldreich65a}
\begin{equation}
Q_{\rm gas}\equiv\frac{\sigma_{\rm gas}\kappa}{\pi G \Sigma_{\rm gas}}\,.
\label{eqn:Q}
\end{equation}
In the thin disk approximation ($kh\ll1$), instability occurs on scales $k$ such that $Q < Q_{\rm crit}\simeq1$ (or equivalently $\omega^2 < 0$ in \Eq{3Ddisp}). A frequently used observable proxy for $Q$ is the ratio of disk circular velocity to root-mean-square velocity dispersion $v_{\rm circ}/\sigma_{\rm disk}$ \citep[e.g.,][]{GarciaBurillo03a, Genzel11a, Kassin12a, Leung19a}.
%

In our stability analysis, we account for the combined effect of gas and stars \citep[derived exactly by][]{Rafikov01a}, and for the non-negligible disk thickness. This is done by adopting an approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modifies the $Q$ parameter for gas and stars by accounting for the vertical velocity dispersion
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars. The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the finite disk thickness reduces the gravity in the vertical direction, thereby making it easier for a system to maintain stability, and thus lowering the critical Toomre $Q_{\rm crit}$ from $\simeq$\,1 to 0.67 \citep{Goldreich65a}.
%
On the other hand, including the contribution of the stellar component promotes gravitational instability, and thus increases
$Q_{\rm crit}$, more-so if the stars have low velocity dispersion.
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

% ------------------
\section{Numerical Simulations}\label{sec:sim}
The simulations used in this work are described by \citet{Pallottini17a, Pallottini17b} and are briefly summarized here.
%
\ncode{Serra}\footnote{Greenhouse in Italian.} is a suite of cosmological zoom-in simulations performed using Eulerian hydrodynamics and adaptive mesh refinement (AMR) techniques to achieve high spatial resolution in regions of interest (e.g., regions of high density).
%
In particular, it uses a modified version of \ncode{ramses} \citep{Teyssier02a} as the AMR backend. The simulation used here covers a comoving box of 20\,Mpc $h$\pmOne in size. The simulation zooms in on a target halo of mass $M_{\rm DM}\simeq10^{11} \Msun$. The Lagrangian region of the halo (2.1\,Mpc $h$\pmOne) has a dark matter mass resolution of $\simeq 6\times 10^4 \Msun$, equivalent at initial density to a baryonic mass resolution of $1.2 \times 10^4 \Msun$. This region is spatially refined with a quasi-Lagrangian criterion based on a mass threshold, so that a cell is refined if its total (dark+baryonic) mass exceeds the mass resolution by a factor of eight. The finest refined cell allowed in the zoom in region has size $l_{\rm cell}\simeq$\,30\,pc (at $z = 6$), i.e. sizes are comparable to the sizes of local giant molecular clouds \citep[e.g.,][]{Sanders85a, Federrath13a, Goodman14a},
%
The models include a non-equilibrium chemical network following e$^{-}$, H$^+$, H$^-$, He, He$^+$, He$^{++}$, H$_2$, and H$_2^+$ \citep{Grassi14a,Bovino16a}. Of particular importance here is that abundances are calculated using an on-the-fly non-equilibrium formation of molecular hydrogen scheme described by \citet{Pallottini17a}.

Star formation is modeled using an H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar
tracks from \ncode{starburst99} and include stellar feedback from supernovae (SNe) as well as winds from OB and asymptotic giant branch stars. Coupling to the gas is implemented via a sub-grid model for blastwaves, which accounts for radiative energy losses inside the cell.
%
The remaining energy is injected into the ISM in both thermal and non-thermal form \citep{Agertz13a}; the latter is a term that mimics unresolved turbulence. The non-thermal energy is not affected by radiative cooling but rather is dissipated on the eddy turn-over time scale \citep{Maclow99a}. See \citet{Pallottini17b} for further details.
%
Radiation pressure on dust and gas is also included. Photoionization from local sources is neglected in this simulation \citep[see][for its effect]{Pallottini19a, Decataldo19a}.

@@

One of the main advantages of studying galaxies in simulations is that we can examine how their dynamical properties evolve with time, which is interesting in order to understand the physical processes that determine the morphology and dynamical properties of galaxies, which in turn affect their \SF.
%
This is advantageous especially at early cosmic epochs, when the densest structures are beginning to form; gas is constantly being accreted onto the central galaxy from the cosmic web and satellite galaxies, thereby leading to bursts of \SF.
%
Meanwhile, tidal forces resulting from interactions with these surrounding galaxies can disrupt the main disk and arms, likely leading to different dynamical states for the molecular structures compared to more evolved galaxies found at a later cosmic time (e.g., some molecular structures may disperse while others may agglomerate into more massive ones).
%
The \SF history\footnote{Note that here the SFR is calculated based on the stellar mass formed in the last 4 Myr within 3.5 kpc from the galaxy center of mass. The SFR plotted in Figure 2 of \citet{Pallottini17b} is a factor of two higher since there the SFR accounts for the contribution from massive satellite galaxies within the virial radius ($\approx$15\,kpc).} of \flower is shown in \Fig{SFH}. The SFR of \flower varies between $\sim$30--80 \Msun\,yr\pmOne as it evolves from an actively accreting phase to a starburst phase after a merger, and then back to a relatively quiescent phase, over the simulated $\approx 700$\,Myr.

In \Sec{dist}, we show the importance of rotation support from large-scale motions in the MCC dynamics 
by comparing their properties in the disturbed phase of \flower and in the disk-like ordered phase.
%
Given the stochastic nature of \flower in its star formation history, we mostly focus on few of its most extreme evolutionary stages in this work (see \Sec{singless}).
%
These phases correspond to (a) an intensely accreting phase and (b) a starburst phase (\Fig{SFH}). We are interested in determining whether MCC properties are sensitive to these different dynamical conditions. For completeness, we also show relationships of MCC properties examined for other evolutionary stages traced in the simulation (see \Sec{ncut}).

@@

In \Fig{h2density}, we show the distribution of H$_2$ number density $n_{\rm H2}$ for \flower during its accreting phase, including the contribution from the gas within 3.5 kpc from the galaxy center.
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of densities where clumps are found based on morphological analysis\footnote{For $n_{\rm H2}\gtrsim1$\,\cc, the fourth Minkowski functional of the H$_{2}$ density field is significantly larger than zero. This implies that the field is made of isolated components. See Fig. 6 in \citet{Pallottini17b}.} \citep{Pallottini17b}.
%
In each evolutionary stage, we identify MCCs by applying density cuts to the {\em H$_2$ density} distribution, i.e., MCCs are selected at $n_{\rm cut} \leq n_{\rm H2}$. We select 10 equally-spaced cuts in H$_2$ density in log scale: ${(n_{\rm cut}/{1 {\rm cm}^{-3})}}\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 18.96, 31.62]$. Note that with these choices we are including MCCs that are not fully molecular ($n = 0.5 n_{\rm H2}$).
